---
title: "Atlas Telecommunications Service Quality Analysis"
output: github_document
---

**Author:** Daniele Melotti<br>
**Date:** Dec 2023

_See README for info about data and list of tasks_.

# 1. Data Visualization and Hypothesis Formulation {#sec1}

We start off by loading the sample data with the repair times of Atlas Telecommunications and summarizing it:
```{r}
atlas <- read.csv(file = "../data/atlascom.csv")

str(atlas)
summary(atlas)
```

We know that we are dealing with a dataframe containing two variables:

* a _Time_ variable, which indicates the repair times,
* a _Group_ variable, which indicates the carrier type; it can be either CLEC or ILEC, but since Verizon is an ILEC, all the entries correspond to ILEC indeed (see [this link](https://www.tutorialspoint.com/difference-between-ilec-and-clec) for differences between CLEC and ILEC).

There are 1664 data entries, the mean value is 8.412 minutes, while the median time is 3.59 minutes.

## Visualize the distribution of Atlas' repair times, marking the mean with a vertical line

Let's see how are the repair times of Atlas telecommunications distributed:
```{r atlas density, fig.align='center'}
# Considering only the Time variable
time <- atlas$Time

plot(density(time), main = "Atlas Repair Times", lwd = 2, col = "tomato")
abline(v = mean(time), lty = "dashed")
```

The plot shows that most of the repair times are contained within 10 minutes or so, however, there are some scenarios in which the repair times increase to roughly 3 hours.

## Formulate the null and alternative hypotheses for testing Atlas' claim about their average repair time

Given that the New York PUC would like to check for any deviations from Atlas' claim regarding the average repair time, we would suppose a *two-tailed scenario*, where the null hypothesis states that the average repair time equals 7.6 minutes, and the alternative hypothesis states that the average repair time is different from 7.6 minutes. Hence:

<div align = 'center'>
  ![](https://latex.codecogs.com/gif.latex?H_0%3A%20%5Cmu%3D7.6)
</div>
<div align = 'center'>
  ![](https://latex.codecogs.com/gif.latex?H_a%3A%20%5Cmu%5Cneq%207.6)
</div>

<br>

# 2. Hypothesis Testing

## Calculate the estimated population mean and construct a 99% confidence interval for it
